% Data analysis.
%%%%%%%%%%%%%%%%

\chapter{Data analysis}



% Introduction.
%%%%%%%%%%%%%%%

\section{Introduction}

This chapter aims to explain not only the steps involved in the data analysis of relaxation data but also how to use the program relax to achieve this.  Although a work in progress, it covers how to calculate the NOE, how to optimise and find the $\Rone$ and $\Rtwo$ relaxation rates, and how to implement model-free analysis.


% Calculating the NOE.
%%%%%%%%%%%%%%%%%%%%%%

\section{Calculating the NOE}


% The sample script.
\subsubsection{The sample script}

The calculation of NOE values is a straight forward and quick procedure which involves two components, the calculation of the value itself and the calculation of the errors.  To understand the steps involved, we will follow in detail the execution of the sample NOE calculation script:

\begin{exampleenv}
\# Script for calculating NOEs. \\
 \\
\# Create the run \\
name = `noe' \\
run.create(name, `noe') \\
 \\
\# Load the sequence from a PDB file. \\
pdb(name, `Ap4Aase\_new\_3.pdb', load\_seq=1) \\
 \\
\# Load the reference spectrum and saturated spectrum peak intensities. \\
noe.read(name, file=`ref.list', spectrum\_type=`ref') \\
noe.read(name, file=`sat.list', spectrum\_type=`sat') \\
 \\
\# Set the errors. \\
noe.error(name, error=3600, spectrum\_type=`ref') \\
noe.error(name, error=3000, spectrum\_type=`sat') \\
 \\
\# Individual residue errors. \\
noe.error(name, error=122000, spectrum\_type=`ref', res\_num=114) \\
noe.error(name, error=8500, spectrum\_type=`sat', res\_num=114) \\
 \\
\# Unselect unresolved residues. \\
unselect.read(name, file=`unresolved') \\
 \\
\# Calculate the NOEs. \\
calc(name) \\
 \\
\# Save the NOEs. \\
value.write(name, param=`noe', file=`noe.out', force=1) \\
 \\
\# Create grace files. \\
grace.write(name, y\_data\_type=`ref', file=`ref.agr', force=1) \\
grace.write(name, y\_data\_type=`sat', file=`sat.agr', force=1) \\
grace.write(name, y\_data\_type=`noe', file=`noe.agr', force=1) \\
 \\
\# View the grace files. \\
grace.view(file=`ref.agr') \\
grace.view(file=`sat.agr') \\
grace.view(file=`noe.agr') \\
 \\
\# Write the results. \\
results.write(name, file=`results', dir=None, force=1) \\
 \\
\# Save the program state. \\
state.save(`save', force=1) \\
\end{exampleenv}


% The run.
\subsubsection{The run}

The script 



% The NOE.
\subsubsection{The NOE}


% Calculating the NOE error.
\subsubsection{Calculating the NOE error}



% Relaxation curve fitting.
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The $\Rone$ and $\Rtwo$ relaxation rates - relaxation curve fitting}
