% The web pages chapter.
%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{The web}



% The relax web sites.
%~~~~~~~~~~~~~~~~~~~~~

\section{The relax web sites}

The main web site for relax is \href{http://nmr-relax.com}{http://nmr-relax.com}.  From these pages, general information about the program, links to the latest documentation, links to the most current software releases, and information about the mailing lists are availible.  There are also search capabilities built into the pages for searching both the HTML version of the manual and the archives of the mailing lists.

The relax website is hosted by Gna! project \href{https://gna.org/}{https://gna.org/} which is described as ``a central point for development, distribution and maintenance of Libre Software (Free Software) projects''.  relax is a registered Gna! project and its primary Gna! webpage, which contains many more technical detailed than the main web site, is \href{https://gna.org/projects/relax}{https://gna.org/projects/relax}.



% The mailing lists.
%~~~~~~~~~~~~~~~~~~~

\section{The mailing lists}
\index{mailing list}

A number of mailing lists have been created covering different aspects of relax.  These include the announcement list, the relax users list, the relax development list, and the relax committers list.

The relax announcement list, "relax-announce at gna.org" is reserved for important announcements about the program.  All new program versions will be announced on this list.  The amount of traffic on this list is relatively low, so if you would like to receive information about relax you can subscribe to the list by vising the information page at \href{https://mail.gna.org/listinfo/relax-announce/}{https://mail.gna.org/listinfo/relax-announce/}.  Previous announcements can be viewed at \href{https://mail.gna.org/public/relax-announce/}{https://mail.gna.org/public/relax-announce/}.

If you would like to ask questions about relax, discuss certain features, receive help, or to communicate on any other subject related to relax, the mailing list "relax-users at gna.org" is the place to post your message.  To subscribe to the list, go to the relax-users information page at \href{https://mail.gna.org/listinfo/relax-users/}{https://mail.gna.org/listinfo/relax-users/}.  You can also browse the mailing list archives at \href{https://mail.gna.org/public/relax-users/}{https://mail.gna.org/public/relax-users/}.

A second mailing list exists for posts relating to the development of relax.  The list is "relax-devel at gna.org" and to subscribe, go to the relax-devel information page at \href{https://mail.gna.org/listinfo/relax-devel/}{https://mail.gna.org/listinfo/relax-devel/}.  Feature requests, program design, or any other posts relating to relax's structure or code should be sent to this list instead.  The mailing list archives can be browsed at \href{https://mail.gna.org/public/relax-devel/}{https://mail.gna.org/public/relax-devel/}.

When replying to a message on these lists, remember to hit 'respond to all' so that the mailing list is included in the CC field.  Otherwise your message will only be sent to the original poster and not back to the list.

One last mailing list is the relax commits list.  This list is reserved for automatically generated posts created by the version control software which looks after the relax source code and these web pages.  If you would like to become a developer, you can subscribe to the list at relax-commits information page \href{https://mail.gna.org/listinfo/relax-commits/}{https://mail.gna.org/listinfo/relax-commits/}. The list can also be browsed at \href{https://mail.gna.org/public/relax-commits/}{https://mail.gna.org/public/relax-commits/}.



% Reporting bugs.
%~~~~~~~~~~~~~~~~

\section{Reporting bugs}
\index{bugs}

One of the philosophies in the construction of relax is that if there is something which is not immediately obvious, then that is considered a design bug.  If any flaws in relax are uncovered, including general design flaws, bugs in the code, or documentation issues, these can be reported within relax's bug tracker system.  The link to submit a bug is \href{https://gna.org/bugs/?group=relax\&func=additem}{https://gna.org/bugs/?group=relax\&func=additem} while the main page for browsing, submitting, viewing the statistics, or searching through the data base is at \href{https://gna.org/bugs/?group=relax}{https://gna.org/bugs/?group=relax}.  Please do not report bugs to personal email addresses or to the mailing lists.

When reporting a bug, please include as much information as posible so that it can be reproduced.  Include information such as the release version or the revision number if the repository sources are being used.  Also include all the steps performed in order to trigger the bug.  Attachment of files is allowed, so scripts and subsets of the input data can be included.  Make sure that you are confident that the problem is trully a bug prior to reporting it, if you have any doubts please feel free to ask on the relax-users mailing list.  To avoid duplicates, be sure that the bug has not already been submitted to the bug tracker.  You can search the bugs from the page \href{https://gna.org/project/search.php?group=relax}{https://gna.org/project/search.php?group=relax}.

Once the bug has been confirmed by one of the relax developers, you may speed up the resolution of the problem by trying to fixing the bug yourself.  If you do wish to play with the source code and try to fix the issue, see the relax development chapter of this manual on how to checkout the latest sources, how to generate a patch (which is just the output of diff in the 'unified' format), and the guidelines for the format of the code.



% Latest sources -- the relax repositories.
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\section{Latest sources -- the relax repositories}

The source code to relax is kept in a version control system called Subversion (\href{http://subversion.tigris.org/}{http://subversion.tigris.org/}).  Subversion, or SVN, allows fine control over the development of the program.  The repository contains all information about every change ever made to the program.  To learn more about the system, the Subversion book located at \href{http://svnbook.red-bean.com/}{http://svnbook.red-bean.com/} is a good place to start.

The contents of the relax repository can be viewed online at \href{http://svn.gna.org/viewcvs/relax/}{http://svn.gna.org/viewcvs/relax/}.  The current sources, assuming that the most recent minor version number is 1.2, can be downloaded using the SVN protocol by typing

\example{\$ svn co svn://svn.gna.org/svn/relax/1.2 relax}

however if this does not work, try the command

\example{\$ svn co http://svn.gna.org/svn/relax/1.2 relax}

to download using the http protocol.  The entire relax repository is backed up daily to \href{http://svn.gna.org/daily/relax.dump.gz}{http://svn.gna.org/daily/relax.dump.gz}.



% News.
%~~~~~~

\section{News}

Summaries of the latest news on relax can be found on the relax web site \href{https://gna.org/projects/relax}{https://gna.org/projects/relax}.  However, more information can be found at the dedicated new page \href{https://gna.org/news/?group=relax}{https://gna.org/news/?group=relax}.



% Downloads.
%~~~~~~~~~~~

\section{Downloads}

The relax distribution archives, the files to download to install relax, can be found at \href{http://download.gna.org/relax/}{http://download.gna.org/relax/}.
